### 2022-07-24 header.tex
add GitHub repo link

% For anonymity
\newcommand{\repo}[1]{https://github.com/mdogucu/teaching-accessibility-manuscript}

% Remove green borders around links
\usepackage{xcolor}
\xdefinecolor{blue}{rgb}{0.28,0.65,0.9}
\hypersetup{
    colorlinks,
    linkcolor={blue},
    citecolor={blue},
    urlcolor={blue}
}