### 2022-11-05 revision_letter.Rmd
some notes in lit review

---
title       : "Teaching Visual Accessibility in Introductory Data Science Classes with Multi-Modal Data Representations"
authors     : "JooYoung Seo & Mine Dogucu"
journal     : "Journal of Data Science"
manuscript  : "JDS2207-022"

class       : "draft"
output      : papaja::revision_letter_pdf
bibliography: ["bib/packages.bib", "bib/references.bib", "bib/teachinga11y.bib"]
---

Dear Dr. Claire McKay Bowen,

Thank you very much for taking the time to consider our manuscript for publication at _`r rmarkdown::metadata$journal`_.
In the following we address your and each reviewers' concerns point-by-point.

# AE

\RC{
    The authors preface their paper with the statement that "In this paper, we argue that accessibility should be taught as early as the introductory course as part of the data science curriculum." This argument is never presented. There is no discussion as to WHY this topic should be included in an introductory course. Why can't it be part of an upper level course?
    }

This is our response

\Assignment{First Author}
\WorkInProgress
\Easy

\RC{
    The authors close with the powerful statement "We believe that an accessible future in data science requires educators to teach accessibility as part of the curriculum deliberately"--but they have failed to present what incorporating it deliberately entails.
    }

This is our response: For example, instructors can include accessibility as one of their evaluation criteria for course assignments or make it as a chance to obtain bonus credits for the students' visualization work.

> This is a section quoted from the revised manuscript to illustrate the change.

\Assignment{Second Author}
\AlmostDone
\Medium

\RC{
    I believe there are two courses of action the authors could take with this paper: a) develop and incorporate a significant argument for accessibility as part of a data science curriculum b) reframe this paper as an experiment in incorporating accessibility methods in an intro course As it currently stands, it falls somewhere in between and falls short of providing a meaningful presentation of either.
    }

This is our response: need to create a section dededicated to a11y, and say that we are presenting this paper as a demonstration how we integrated a11y into our courses.

\Assignment{First Author}
\Done
\TimeConsuming
\Hard
\NotEstimated

# Reviewer

\RC{
    What are the significances of teaching visual accessibility at the 'introductory' data science level (as opposed to introducing them at a higher level of data science courses after students have a solid foundation in the discipline?
}

This is our response

\Assignment{First Author}
\WorkInProgress
\Easy

\RC{
If the visual aid accessibility was introduced in the 'introductory' courses, will students see it again in higher-level courses? In other words, would other faculty members who teach higher levels continue to use this in their courses? Students would not retain the information if they were only to see visual aid accessibility at the beginning of their data science journey.
}

This is our response: good point so it is critical to share the same principle across different and with upper level courses .

\Assignment{First Author}
\WorkInProgress
\Easy

\RC{
As visual aid accessibility is an essential part of various disciplines and career paths (science, engineering, business, arts and letters, etc.), would it be more beneficial to have a separate course (as a university-core or university-elective course) teaching visual aid accessibility rather than to integrate it into the foundational data science courses?
}

According to @shinoharaWhoTeachesAccessibility2018,

> The most critical barriers to teaching accessibility that faculty reported were the absence of clear and discipline-specific accessibility learning objectives and the lack of faculty knowledge about accessibility.

> Faculty desired resources that were specific to the areas of computing in which they teach rather than general accessibility resources and guidelines.

\Assignment{First Author}
\WorkInProgress
\Easy

\newpage

# References

::: {#refs custom-style="Bibliography"}
:::