### 2022-11-08 manuscript.Rmd
define AT

---
title: |
  Teaching Visual Accessibility in Introductory Data Science Classes with Multi-Modal Data Representations
shorttitle: Teaching Accessibility in Data Science
author:
- name: JooYoung Seo
  affiliation: "1"  # Must be quoted when including comma.
  email: "jseo1005@illinois.edu"  # Only corresponding authors require email.
- name: Mine Dogucu
  affiliation: "2,3"

affiliation:
- id: 1
  institution: "School of Information Sciences, University of Illinois Urbana-Champaign, 614 E. Daniel St., Champaign, IL 61820, USA"
- id: 2
  institution: "Department of Statistical Science, University College London"
- id: 3
  institution: "Department of Statistics, University of California Irvine"

abstract: |
  Although there are various ways to represent data patterns and models, visualization has been primarily taught in many data science courses for its efficiency. Such vision-dependent output may cause critical barriers against those who are blind and visually impaired and people with learning disabilities. We argue that instructors need to teach multiple data representation methods so that all students can produce data products that are more accessible. In this paper, we argue that accessibility should be taught as early as the introductory course as part of the data science curriculum so that regardless of whether learners major in data science or not, they can have foundational exposure to accessibility. As data science educators who teach accessibility as part of our lower-division courses in two different institutions, we share specific examples that can be utilized by other data science instructors.
keywords: 
  - Data representations
  - Curriculum
  - R

# the following edits should be done by Journal typesetters
page: 1  # set the first page number
month: July
year: 2020
volume: xx
issue: xx
doi: "xx.xxxx/xxxxxxxxx"
received: "April, 2020"  # optional: comment it out if no received date
accepted: "May, 2020"  # optional: comment it out if no accepted date

bibliography: ["bib/references.bib", "bib/packages.bib", "bib/teachinga11y.bib"]
csl: "bib/chicago-author-date.csl"
output:
  rticles::jds_article: 
    citation_package: default
    includes:
      in_header: header.tex #header-hidden.tex can be used when creating an anonymized version for review
  bookdown::pdf_book:
    base_format: rticles::jds_article # for using bookdown features like \@ref()
    citation_package: default
---

```{r setup, include=FALSE, echo=FALSE, purl=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.path = "figure/")

# deactivate addition of preamble when this format is used with bookdown
options(bookdown.theorem.preamble = FALSE)
```

```{r pkg, include=FALSE, echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE, purl=FALSE}
# Load required packages
## Any packages loaded in this chunk will automatically be added to `bib/packages.bib` for citation use
library(ggplot2)
library(scales)
library(BrailleR)
library(sonify)
library(tactileR)
library(tidyverse)
library(shiny)
library(knitr)
library(gt)
library(colorblindr)
library(palmerpenguins)
```

# Introduction

According to LinkedIn's U.S. Emerging Jobs Report, data scientists rank among the top emerging jobs [-@linkedin]. With the urgent need for training a higher number of skilled data scientists, many institutes of higher education are developing their own data science curricula. However, a consensus on what should be included in such curricula has not yet been reached. One of the few guidelines on the topic is written by the ACM Data Science Task Force and titled *Computing Competencies for Undergraduate Data Science Curricula*. In this report, accessibility is a "foundational consideration" at the knowledge level when it comes to displaying data [@danyluk2021computing].

In fact, training on accessibility has been increasingly becoming an important job-readiness aspect in computing and information sciences [@shinoharaWhoTeachesAccessibility2018; @kawasTeachingAccessibilityDesign2019]. For instance, some notable big tech companies, such as Yahoo!, Facebook, Microsoft, Google, and Adobe, have been leading a TeachAccess (teachaccess.org) Consortia to represent this increased demand for accessibility knowledge and to increase the number of computer and information science faculty teaching accessibility [-@shinoharaWhoTeachesAccessibility2018]. The Partnership on Employment & Accessible Technology (PEAT), in their Accessible Technology Skills Gap Report [-@peat], also highlights this trend--84% of industry correspondents that they work with say "it is important or very important for them to hire developers and designers with accessible technology skills." The growing interest in and importance of teaching accessibility can also be found in professional education societies. For example, accessibility has been recently added to engineering design accreditation requirements by Accreditation Board for Engineering and Technology [ABET; abet.org] [@shinoharaWhoTeachesAccessibility2018].

Even though this statistic may be different specifically for data scientists, data science as a highly interdisciplinary field across computing and information sciences also shares the rising demand for accessibility preparedness. Furthermore, data science is broad with the intersection in careers such as developers (e.g., data dashboard and package developers) and designers (e.g., data visualization experts). Therefore, the impact of teaching and learning accessibility is as far-reaching as its wide scope.

Data science consists of an iterative process of importing, tidying, transforming, visualizing, and modeling data [@wickham2016r]. While each stage has unique challenges, data visualization may pose fundamental barriers against people with permanent, situational, and temporary disabilities (e.g., visual, cognitive, and motor impairments, to name a few) in accessing information [@marriottInclusiveDataVisualization2021; @leeReachingBroaderAudiences2020]. For example, people with cognitive and learning disabilities often have difficulties understanding the abstractions and symbolic conventions used in visualizations, and people lacking fine motor coordination or having tremors may be unable to operate interactive visualizations, not to mention people who are blind and have low vision facing extra challenges in visualization access [@marriottInclusiveDataVisualization2021]. @marriottInclusiveDataVisualization2021, in their call to action paper, highlights the serious impact of inaccessible visualizations: "This is an equity issue with severe consequences, because a lack of access to visualizations and underlying data impacts health, educational and work opportunities, and lifestyle." <!-- Need to find a specific page number for this quote. -->

Despite the increasing demand for accessibility knowledge in both industries and academia and the critical impact of accessible visualizations, accessibility has not yet found its place in the data science course and teaching materials. One main reason for this could be that accessibility is often incorporated into human-computer interaction (HCI) and design courses [@shinoharaHowDesigningPeople2016; @shinoharaTeachingAccessibilityTechnology2017a], introductory web development [@koAccessComputingPromotesTeaching2016; @rosmaitaAccessibilityFirstNew2006], software engineering courses [@martin-escalonaIntroductionTopicAccessibility2013; @ludiIntroducingAccessibilityRequirements2007; @ludiAccessEveryoneIntroducing2002] or the digital arts [@barata2019inclusion], if at all covered. However, data science is an interdisciplinary field that is taught in various disciplines, including not only computer and information sciences but also statistics, business schools, political science, biological sciences, and more [@schwab2021data]. Thus, we posit that the resources and conversations around accessibility have not yet extended to the broader data science education community at the scale that they should. For instance, many introductory data science courses [@yan2019first; @baumer2015data], textbooks [@adhikari2019computational; @wickham2016r], or curricula [@schwab2021data] either fail to mention accessibility at all or only focus on the accessing data.

In response to the need to develop accessible teaching materials within the data science discipline, this paper demonstrates and shares some teaching strategies that we have employed in our teaching introductory data science courses. In this manuscript, we argue that accessibility should be taught early in the data science curriculum so that regardless of whether learners major in data science or not, they can have foundational exposure to accessibility.

Some people may wonder why we need to teach accessibility as part of an introductory data science course as opposed to covering accessibility in a separate and general course module. According to the recent faculty survey on teaching accessibility [@shinoharaWhoTeachesAccessibility2018], the most critical barriers to teaching accessibility reported by 1,857 faculty from 318 institutions were "the absence of clear and *discipline-specific accessibility* [emphasis added] learning objectives and the lack of faculty knowledge about accessibility." <!-- Need to find the exact page number for this quote -->
This means faculty often have trouble making sense of how they can integrate accessibility into their specific teaching domains, which requires applied knowledge of accessibility beyond general concepts.

We share our experiences and strategies in the hope that our discipline-specific (i.e., data science and data visualization) accessibility teaching cases can be used as scaffolding guides for other data science instructors to build upon.

Another critique we may receive includes the timing of when to introduce accessibility concepts in data science courses. For example, some instructors may prefer to put teaching accessibility on the back burner so that they can save more time for their students to learn data science fundamentals first and adopt accessibility principles in an upper-level course later. While the decision is subject to each institution's and individual instructor's discussion, we believe the sooner, the better. This is because: (1) some students may take an introductory level only, which means they will lose a chance to get exposed to this demanding knowledge unless they elect the upper-level course covering accessibility; (2) this kind of separation may unintentionally contribute to misconceptions, such as accessibility is difficult and/or accessibility is optional, not essential. When properly incorporated, accessibility can be taught along with other data science fundamentals (e.g., wrangling, merging, tidying, visualizing) while not necessarily putting more cognitive burdens against students' current learning. Rather, students can take a chance to become creative design thinkers by making their data science work more meaningful and accessible through "empathy learning" [@putnamTeachingAccessibilityLearning2015]. Thus, we suggest teaching accessibility as part of an introductory data science course for students to practice coupling their data science knowledge with accessibility principles as early as possible.

As data science educators who teach accessibility as part of our lower-division courses in two different institutions, we share specific examples that can be utilized by other data science instructors. We teach R [@R-base] as the main language in our courses; thus, our examples are shaped around using R. However, instructors using a different set of language(s) or even teaching language-agnostic courses may still find the content applicable to their own courses.

# Learning Goals and Objectives

Although there are various ways to represent data patterns and models, visualization has been primarily taught in many data science courses for its efficiency [@kimAccessibleVisualizationDesign2021]. 
Such vision-dependent output may cause critical barriers against not only blind and visually impaired people but also  people with cognitive and learning impairments and people having motor impairments [@kimAccessibleVisualizationDesign2021; @marriottInclusiveDataVisualization2021; @leeReachingBroaderAudiences2020]. 
We argue that instructors need to teach multiple data representation methods so that all students can produce data products that are more accessible. 
To this end, we have considered multi-data representations and their relation to accessibility in our courses, and that's what we will demonstrate throughout this paper.

In incorporating accessibility to our course contents, we also wanted our students to be familiar with assistive technologies, and since our focus has been on visual accessibility, it was a natural fit to include screen readers. 
Screen reader is an assistive technology that supports blind or visually impaired people in using their computer by reading aloud the contents. Recently, screen reader has been provided as one of the built-in accessibility features on operating systems, such as Windows Narrator and Apple's VoiceOver.

However, we also want to make our teaching as inclusive as possible to other types of dis/abilities so that we are not just confined to issues of visual disabilities.

Therefore, our curricular goal is to broadly ensure that the current and the next generation of data scientists provide public-facing outputs, including websites and analysis reports, in accessible forms and representations for people with diverse abilities. To achieve this goal, we set three specific and minimal learning objectives like below.

Students should:

1. get familiar with Americans with Disabilities Act;
2. use at least one assistive technology;
3. consider different representations of data.

The first objective provides our students with a chance to learn legislative concepts and empathy building concerning diversity issues. As both the authors teach data science in the United States, it was a natural choice for us to make students aware of the Americans with Disabilities Act (ADA). 
The ADA is a civil rights law that prohibits discrimination against individuals with disabilities. 
Regardless of their own disability status, as future data scientists, we expect our students to produce data science products that are accessible to all, not only because it is the right thing to do but also because they may have future responsibilities as data scientists to comply to ADA.

The second learning objective echoes our argument on accessible data representation beyond one-type modality. People with disabilities may rely on different types of assistive technology in their day-to-day lives. 
Assistive technology (AT) is defined as "products, equipment, and systems that enhance learning, working, and daily living for persons with disabilities" (Assistive Technology Industry Association; [https://bit.ly/3fJN2aJ](https://bit.ly/3fJN2aJ)). Examples of assistive technology include, but are not limited to, walking sticks, wheelchairs, screen readers, foot-activated switches, joysticks, sip and puff
controllers, eye tracking device, and speech recognition input. As we noted above, we introduced screen reader as one of these examples when our students were working on a data visualization project; however, we acknowledge that various assistive technologies, such as speech recognition and video captioning, could be integrated into data science courses in diverse contexts (e.g., data collection and data presentation contexts).

It is important to note that the first two objectives are not data science specific but are related to accessibility on a broader scale. 
The third objective, however, is data science specific. In the remaining section of this paper, we will focus further on the third objective by illustrating how we have taught multiple data representations.

## Data Visualization

Data visualization is the most common representation of data. Despite their use by many, data visualizations are not accessible to all. 
Colors are an important part of data visualizations and an aspect that can limit accessibility for color-blind people. 
When teaching data visualizations, we include three basic rules for incorporating colors into data visualization. Students should be able to

1. simulate color blindness (if they are sighted and not color blind);
2. pick color blind friendly colors;
3. not only rely on color for differentiating data.

We show an example of incorporating these three basic rules by using the Palmer Penguins data [@R-palmerpenguins]. 
We visualize the relationship between flipper length and bill length of penguins and color the data points differently for each species in Figure \ref{fig:penguins-basic}.

```{r penguins-basic, echo=FALSE, fig.cap="Sample scatterplot showing the relationship between flipper length (mm) and bill length (mm) of penguins for three different species of penguins", fig.alt="Sample scatterplot showing the relationship between flipper length in mm on the x axis and bill length in mm on the y axis. Flipper lengths vary from about 170 to 230 and bill lengths vary from about 35 to 60. Overall there is a moderate positive relationship. Each data point is colored differently for three species as Adelie, Chindtrap and Gentoo. Adelie have short flipper length and short bill length in comparison. Chinstrap have low flipper length but high bill length. Gentoo have high flipper length and high bill length. Data points are in red, green, and blue for Adelie, Chinstrap and Gentoo penguins respectively.", warning=FALSE, fig.height=3}
library(ggplot2) # It seems like the pkg chunk is not loading packages. I did not want to make changes there since I do not know what your intention was.
library(palmerpenguins)

fig <- ggplot(penguins, aes(flipper_length_mm, bill_length_mm, color = species)) +
  geom_point() +
  labs(
    x = "Flipper Length(mm)",
    y = "Bill Length (mm)",
    color = "Species"
  ) +
  theme_bw()

fig
```

The `colorblindr` package [@R-colorblindr] provides a set of functions that can be useful in teaching and practicing color blindness accessibility. 
For instance, `colorblindr::cvd_grid()` function provides color-deficiency simulations of a given plot. 
In the code chunk below, the plot is named as `fig`. 
The code creates a grid of four plots where each plot shows a different color vision deficiency, including deuteranomaly, protanomaly, tritanopia, and desaturated.
This grid allows those without any color vision deficiencies to experience what data visualizations would look like to a sighted person that has color defiency.

```{r penguins-cvd, fig.cap="Color Blindness Simulation with colorblindr", fig.alt="A 2 by 2 grid of four scatterplots. The scatterplots are the same as the scatterplot in the previous figure except for colors. From left to right and top to bottom the plots read deutanomly, protanomly, tritanomly, and desaturated. The deutanomly plot shows two sets of points that are close to brownish green and one set of points that are violet. The protanomly plot is similar in colors, one of the brownish green is closer to gray. In the tritanomly plot, it is difficult to distinguish the green/blue colors. There is also a pink color that stands out. In the desaturated plot all the points are gray and the different species are indistinguishable.", warning=FALSE, echo=TRUE}
colorblindr::cvd_grid(fig)
```

For their data visualizations, instead of relying on the default color selection of R packages, students are introduced to the Okabe-Ito color palette, which is known to be accessible to people with color vision deficiencies [@okabe]. 
One can find specific colors in the pallette by the R code `palette.colors(palette = "Okabe-Ito")`.
In addition, students can be introduced to `colorblindr::scale_color_OkabeIto()` and `colorblindr::scale_fill_OkabeIto()` to automatically switch to using this color pallette by adding a layer to a ggplot presentation. 
For instance, the plot in Figure \ref{fig:penguins-basic} can be switched to the Okabe-Ito color palette as shown in Figure \ref{fig:penguins-okabe} by using the code below.

```{r penguins-okabe, fig.cap="Scatterplot Using Okabe-Ito Color Pallette", fig.alt="The scatterplots are the same as the scatterplot in the original penguins plot except for colors. Data points are in orange, sky blue, and bluish green for Adelie, Chinstrap and Gentoo penguins respectively.", warning=FALSE, fig.height=3, echo=TRUE}
fig + colorblindr::scale_color_OkabeIto()
```

Last but not least, students should learn not to rely only on color to differentiate data points, in this case, Species. 
For instance, they can be introduced to shapes and have data points that are circles, triangles, and squares for each species. 
This can also be extended to line graphs as having continuous or dashed lines.
Facetting can also be utilized. 
For instance, each of the species could have its own scatterplot side-by-side.

In this manuscript and in our introductory courses, we focus on static data visualizations. 
Those who teach interactive data visualizations may also consider covering cognitive and motor accessibility. 
Our focus here is on visual accessibility, but it is worth noting that all subdomains of accessibility are relevant to data science learners and can be incorporated into the curriculum as seen fit.

## Data verbalization (i.e., alt text)

Our definition of "data verbalization" refers to a way of representing data patterns using verbal description with alternative text (i.e., alt text). 
Images without alt text markup are not readily accessible to assistive technologies (i.e., screen readers; refreshable braille displays; text-based voice commands) and web search parsers. 
However, appropriately designed alt text can not only give assistive technology users minimum access to visualized data but also provide all students with a much richer context [@lundgardAccessibleVisualizationNatural2022].

There are two ways to create alt text: (1) auto-generated alt text; and (2) manual alt text. Currently, `BrailleR::VI()` [@R-BrailleR] function has a capability to auto-generate alt text for basic R graphics and ggplot objects (see Figure \ref{fig:VI-example}). 
Some of the supported graphic types include: `graphics::hist()`, `graphics::boxplot()`, `ggplot2::geom_bar()`, `ggplot2::geom_histogram()`, and `ggplot2::geom_boxplot()`.

```{r VI-example, fig.cap = "Sample histogram for automatic data verbalization.", fig.alt = "Bar chart of penguins species with its verbal description automatically generated by `BrailleR::VI()` function.", echo=TRUE, include=TRUE, warning=FALSE, cache=TRUE, fig.height=3}
library(ggplot2)
library(BrailleR)
library(palmerpenguins)

# Create graph
g_species <- ggplot(data = penguins, mapping = aes(x = species)) +
  geom_bar()

g_species

# Obtain auto verbalization
# (Added feature in BrailleR >= 0.32.1): `BrailleR::VI()` function is automatically called against supported graphics if BrailleR package is currently loaded in R session
# BrailleR::VI(g_species)
```

Even though the automated alternate text has some use with its limitations, it does not often directly convey the message a visualization is displaying.
Thus, it is important to write alternate text manually and teach students how to write one.
We rely on Cesal's “simple formula for writing alt text for data visualization” [-@cesal], which is as follows: “alt= _Chart type_ of _type of data_ where _reason for including chart_.” 
She also recommends the inclusion of a link to data somewhere within the text. 
Similarly, for a meaningful alternate text, @canelon suggest that the description should convey the meaning of the data, variables on the axes, scale, and type of plot should be included.

Given these recommendations, we provide the alternate text we wrote for Figure  \ref{fig:penguins-basic}. "Sample scatterplot showing the relationship between flipper length in mm on the x-axis and bill length in mm on the y-axis. Flipper lengths vary from about 170 to 230, and bill lengths vary from about 35 to 60. Overall there is a moderate positive relationship. Each data point is colored differently for three species as Adelie, Chinstrap, and Gentoo. Adelie has a short flipper length and short bill length in comparison. Chinstrap has low flipper length but high bill length. Gentoo has a high flipper length and high bill length. Data points are in red, green, and blue for Adelie, Chinstrap, and Gentoo penguins, respectively." 
Note that we specify colors as they would be relevant as readers switch from Figure  \ref{fig:penguins-basic} to Figure  \ref{fig:penguins-okabe}.

Once alt text is prepared either programmatically or manually, it can be embedded in static and dynamic plot objects in R, such as `ggplot2::labs(..., alt = "[alt text]")`, 
`shiny::renderPlot(..., alt = "[alt tex]")` [@R-shiny]. We also ask our students to provide alt text within their R Markdown chunks by using `fig.alt = "[alt text]"` `knitr` chunk option [@knitr2014], and within Pandoc-flavor/Common Markdown with `![alt text here.](path/to/file)` syntax [@macfarlanePandoc2022].

## Data sonification using sound

Our definition of sonification refers to a data representation method using stereo sound and various audible patterns. For example, Figure \ref{fig:plot-xy} can be made audible by using `sonify` package [@R-sonify] like below. The output is audio which can also be accessed as an audio file at \url{\audio{}}.

```{r sonify-x-y, echo=TRUE, include=TRUE, cache=TRUE}
# Representing the relationship between the two numerical variables via stereo sound
library(sonify())

# Prepare variables
x <- 1:5
y <- 1:5

sonify(x, y)
```

In this example, a scatter plot is represented by sound. Position on the X axis is communicated using the left-to-right stereo effect; the pitch of the sound indicates the position on the Y axis.

Data sonification is very useful when representing linear patterns; however, there is much room for improvement in terms of complex data sonification as of this writing. Thus, we suggest using sonification for simple linear regression models.

```{r plot-xy, fig.cap="Visualizing the relationship of two numerical variables with scatter plot.", fig.alt="Scatter plot representing the positive relationship between x and y variables. There are five points on the plot at x,y values as 1,1 - 2,2 - 3,3 - 4,4 - 5,5", echo=FALSE, fig.height=5}
# Prepare variables
x <- 1:5
y <- 1:5

# Representing the relationship between x and y variables via plot visualization
plot(x, y)
```

## Data tactualization using Swell Form machine

Data tactulization refers to making data visualization in a tactile form so that it can be touchable by fingers. 
Although this requires some expensive embossing hardware (e.g., braille embosser or Swell Form Heating Machine), this may be one of the most effective and accessible data representations for those who are blind or visually impaired. 
The `tactileR` [@R-tactileR] package in R can generate a ready-to-emboss pdf file from R graphics and ggplot objects. 
Students and instructors can produce their own tactile graphics by using the following functions: `tactileR::brl_begin()`; [their graphic object]; `tactileR::brl_end()` (see Figure \ref{fig:tactile-example}). 
For those who do not have access to embossing hardware, the learning process on tactualization can be supported with a video that shows the process of printing a tactile graph available at \url{\video{}}.

```{r tactile-producing-code, echo=TRUE, include=TRUE, cache=TRUE, message=FALSE, results="hide"}
library(palmerpenguins)
library(tactileR)

brl_begin("figure/tactile-boxplot.pdf")
with(penguins, boxplot(body_mass_g ~ species))
brl_end()
```

```{r tactile-example, echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE, dependson="tactile-producing-code", fig.cap="Ready-to-print tactile graph in braille.", fig.alt="Boxplot showing the different body mass between penguins species. The plot has three boxes for each Penguin species. On the x and y-axes, labels and values are displayed in braille.", , fig.align = "center", out.height="50%"}
knitr::include_graphics("figure/tactile-boxplot.pdf")
```

# Delivery of Content

We believe that a natural connection between data science and accessibility can be achieved in the data science classroom while presenting data visualization and then introducing the other aforementioned forms of data representation.
While delivering the content, instructors should try to provide as many opportunities as possible for students to experience using these tools themselves. 
For instance, each student should use a screen reader at least once regardless of their visual ability status and hear the alternate text that they have written.

Ample opportunities in learning accessibility should also include assessments. 
For instance, instructors incorporating accessibility should modify assessment instructions and rubrics to include accessibility. 
This will prevent students from just hearing about accessibility in the lecture and then forgetting about it. 
Ideally, the assessments should not treat accessibility as a learning objective at a single point in the academic term. 
For instance, if accessibility is covered in Week 3 of the term, students can (and should) still be expected to write alternate texts for visualizations in Week 10.

In this manuscript, we have focused on domain-specific (i.e., data science) knowledge of accessibility using a particular language (R). 
Depending on the courses taught, instructors may want to extend the accessibility content to non-data contexts. 
In addition, they may adopt the content in their courses with other languages such as Python and SAS or even in their language-agnostic courses. 
Overall, if taught well, a student learning accessibility content, for instance, should be able to include alternate text in their PowerPoint presentation in their Anthropology class. 
In other words, accessibility knowledge should extend beyond data science and beyond any tool.

Although we have mainly focused on tools and different representations of data to have students develop accessible content, a more crucial accessibility feature is reproducible workflows. 
Scholars often focus on reproducibility and replication of scientific findings. 
We believe reproducibility can often serve as an accessibility feature in a data science context. 
A reproducible open-source workflow allows users who are blind or have a visual impairment to examine data in their preferred form of data representation.
Thus, in addition to teaching different representations of data, instructors may also consider teaching reproducibility.

# Closing Remarks

We believe that an accessible future in data science requires educators to teach accessibility as part of the curriculum deliberately. 
We also acknowledge that what we have presented here is only a small portion of the intersection of visual accessibility and data science. 
The second co-author was supported through a Teach Access grant to develop curricular materials on teaching accessibility as part of the data science curriculum. 
Educators interested in teaching accessibility can follow the organization's work at https://teachaccess.org/. 
The reproducible R source code of this manuscript can be found on a GitHub repository at \url{\repo{}}.

# References {-}

```{r pkg_bib, include=FALSE, echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE, dependson="pkg"}
knitr::write_bib(file = "bib/packages.bib")
```

# Lit Review

## @shinoharaWhoTeachesAccessibility2018

### RP

Industry demand for software developers with knowledge of accessibility has increased substantially in recent years. However, there is little knowledge about the prevalence of higher education teaching about accessibility or faculty's perceived barriers to teaching accessibility.

### RQ

* RQ1. Who is teaching accessibility?
* RQ2. What barriers do faculty see to teaching accessibility?
* RQ3. What factors predict who is teaching accessibility?

### DC

surveyed 14,176 computing and information science faculty in the United States. We received a representative sample of at least one response from 318 of the 352 institutions we surveyed, totaling 1,857 responses.

### Findings

175 institutions (50%) had at least one instructor teaching accessibility and that no fewer than 2.5% of faculty overall teach accessibility. Faculty that teach accessibility are twice as likely to be female, to have expertise in HCI and software engineering, and to know people with disabilities.

Overall, 20% of faculty (375) responded “Yes” to the prompt “Do you teach courses that incorporate topics about accessibility?”

The most critical barriers to teaching accessibility that faculty reported were the absence of clear and discipline-specific accessibility learning objectives and the lack of faculty knowledge about accessibility.

Faculty desired resources that were specific to the areas of computing in which they teach rather than general accessibility resources and guidelines.

### Quotes

<!-- Increasing interest and importance of accessibility -->
> Consortia like TeachAccess (teachaccess.org), led by Yahoo!, Facebook, Microsoft, Google, and Adobe, represent this increased demand for accessibility knowledge, and are leading efforts to increase the number of computer and information science faculty teaching accessibility.

> In parallel, professional societies such as ABET [abet.org] have recently added accessibility into engineering design accreditation requirements.

<!-- Examples of how accessibility is integrated into computing and information sciences courses -->
> some have incorporated accessibility in software engineering courses [@martin-escalonaIntroductionTopicAccessibility2013; @ludiIntroducingAccessibilityRequirements2007; @ludiAccessEveryoneIntroducing2002], as part of introducing web development [@koAccessComputingPromotesTeaching2016; @rosmaitaAccessibilityFirstNew2006], or in HCI and design courses as a central component of design process [@shinoharaHowDesigningPeople2016; @shinoharaTeachingAccessibilityTechnology2017a].

<!-- Two efforts are required to make broader impacts for teaching accessibility -->
> Incorporating instruction on accessibility at scale requires two major elements: (1) methods for how to teach accessibility in computer science (CS) and related fields, and (2) large scale adoption and implementation of these methods by faculty.
<!-- In conclusion, we have just demonstrated some methods to make data science visualization more accessible, focusing on the second element. However, we need more broad effort to make it at a large scale as @shinoharaWhoTeachesAccessibility2018 said. -->

> These works suggest that large scale curricular change is not a matter of top-down decree, but one of fostering local change agents who are intimately aware of the culture of specific institutions and who can devise and implement strategies to work within them.

<!-- Active culture movement is required towards teaching accessibility -->
> theories of curricular change argue that change only comes through social and instructional discourse among faculty organically contributing to, and constructing, instructors’ knowledge [@barnettConceptualisingCurriculumChange2001; @kirkTeacherVoiceOwnership2001].

> Kezar and Eckel, in an examination of several institutions, determined that institutional culture could enhance or stymie curriculum changes [@kezarEffectInstitutionalCulture2002].

## @kawasTeachingAccessibilityDesign2019

### RP

Most CS students learn little about accessibility in higher education; this is partly because most CS faculty know little about accessibility.

higher education CS faculty lack a model of professional development for learning to teach new topics.

### DC

conducted 18 semi-structured interviews with U.S. CS faculty, asking them to explore a prototype of a web-based professional development tool that linked accessibility topics to CS topics.

we investigated the feasibility of a "micro" professional development model for teaching accessibility in CS courses that could be used at scale.

### Findings

many organizational factors limited faculty's autonomy to integrate accessibility in many of their courses.

individual values and knowledge constrained faculty's ability and willingness to both learn and integrate accessibility topics into their courses.

However, many faculty expressed desire to teach accessibility in their courses if they had access to even basic accessibility content and materials to use in their courses.

## @marriottInclusiveDataVisualization2021

### Quotes

<!-- how many people with disabilities in the world? -->
> A 2011 report from the WHO (https://bit.ly/3raT0kg) estimates that more than 1 billion people—15 percent of the world’s population— live with some form of disability, of whom about one-fifth experience significant difficulties in functioning. While it may be obvious that people who are blind or have low vision (BLV) would face challenges in accessing data visualizations, this also holds for other kinds of disability. People who have cognitive and learning disabilities (CALD) frequently have trouble making sense of the abstractions and symbolic conventions used in visualizations, and people who lack fine motor coordination or have tremors may be unable to operate interactive visualizations.

<!-- Three types of major disabilities  affected by inaccessible viz -->
> There are three main kinds of disability that may affect access to visualizations: visual, cognitive, and motor impairments. It is not uncommon for people to have more than one kind of impairment.

<!-- Why is accessible Visualization important? -->
> this is an equity issue with severe consequences, because a lack of access to visualizations and underlying data impacts health, educational and work opportunities, and lifestyle. For instance, the BLV community has identified a lack of access to educational materials [@butlerUnderstandingGraphicalChallenges2017] and online data visualizations about Covid-19 as a significant concern (https://bit.ly/2NJzy0C).

<!-- People are not paying enough attention to accessible viz -->
> The visualization research community, however, has not paid enough attention to the needs of people with disabilities [@leeReachingBroaderAudiences2020].

<!-- Accessible viz for all beyond disabilities -->
> Inclusive approaches to visualizations are likely to help more than those people with permanent disabilities. There are also hundreds of millions of people with temporary or situational disabilities. New, inclusive approaches could readily become “informational curb cuts” that benefit everyone.

<!-- Tactile representation for blind people -->
> The Braille Authority of North America (BANA; http://www. brailleauthority.org) recommends the use of tactile graphics to show spatial data. These are usually raised line drawings, but more recently, 3D-printed models have been used. Research suggests that despite the differences between tactile and visual perception, spatial perception and reasoning is fundamental to the use of both visual and tactile graphics, and that tactile maps, scatter plots, and network diagrams offer similar benefits to the corresponding visual graphic (e.g., @yangTactilePresentationNetwork2020a).

<!-- Tactile production is expensive -->
> Unfortunately, the cost and time for the production of tactile graphics limits their use for data exploration. More recently, researchers have explored the multimodal presentation of graphics on touchscreen devices, providing audio and haptic feedback when the reader’s fingers touch the graphic. There is considerable opportunity to explore data-exploration tools leveraging refreshable tactile displays as well as touchscreen-based presentations, as these may provide similar benefits to tactile graphics.

<!-- critical disability studies -->
@mankoffDisabilityStudiesSource2010

<!-- The following is good to review when having more time -->

## ## @joynerVisualizationAccessibilityWild2022

### rp

### rq

### dc

### ### findings

## @koAccessComputingPromotesTeaching2016

### RP

### RQ

### DC

### Findings